---
title: "Creating and Working with Vectors"
author: "kimberly yom"
date: "2022-08-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

When you make an R markdown file, the title you give it is different from the file name.

```{r}
library(tidyverse)
#tells us the packages it's loading
library(palmerpenguins)
```

## Let's make some vectors!

```{r}
marmots <- c("blue", "green", 4, "yellow")

#I can now see the object "marmots" in the environment window
# in vectors, every element needs to be the same class, ex. characters vs. integers, so even though 4 is a number not in quotes, it'll get processed as a character bc chr are the most universal, since R makes assumptions on what class you want.
```

Let's check the class

```{r}
class(marmots)
```

```{r}
pika <- c(12.4, 6.8, 2.9, 8.8, 8.5)
class(pika)
# scalar multiplier

scalar_pika <- 5.2 * pika
#will return every element in the vector multiplied by 5.2
```

When you want the values of an object, you have to call for it in your code, ex. I have to type scalar_pika in a code chunk to actually have the results show up when I knit

```{r}
bananas <- c(1, 2, 3)
apples <- c(6, 7, 8)
# can do math with them since they're equal lengths

bananas + apples

bananas - apples

bananas %*% apples
#%*% gives you the dot product
```

### Matrices - how to make them

```{r}
my_values <- seq(from = 1, to = 10)
#if you don't put a "by" value for the seq() command, it'll assume you're incrementing by 1

#if instead you want to tell it the quantity of values you want, use length

#my_values_length_ex <- seq(from = 1, to = 10, length = 50)
#my_values_length_ex

#this will produce 50 equally spaced values starting from your from number and going to your to number

my_matrix <- matrix(data = my_values, nrow = 2, ncol = 5, byrow = TRUE)
my_matrix
#data, how many rows i want, how many columns i want, byrow = TRUE fills first row then moves to second row once it's filled, read matrix under Help tab

#if number of values doesn't match up with number of rows and columns, will produce error message

5*my_matrix

#make sure to regularly check the class as a troubleshooting starting point by class(object)

class(my_matrix)
```

```{r}
glimpse(penguins)
#shows previews of the rows and can see the annotated versions of classes of my row, ex. <fct> is factor
```

The purpose of terminal is universal language that runs commands without the use of a development environment. You can work in the command line for any environment and save your work to git

Working in console vs. terminal: Your console is giving commands to your DE, terminal is giving commands to your computer

## Terminal commands (commands are case-sensitive):
Input these commands in the terminal tab

**pwd** = print working directory - should tell you which directory you're working in

**git status** = tells us the modified files that haven't been staged

**git add -A** = stages all modified files (equivalent of checking the Staged boxes of every modified file) (doesn't look like it - it won't actually check the Staged boxes)

**git commit -m "commit message"** = commits the files with the commit message in the quotes

**git push** = pushes files to the remote repo
